---
title: "LaTeX2Exp"
author: "Stefano Meschiari"
output: html_document
---
# LaTeX2Exp
**latex2exp** is an R function that parses and converts LaTeX math formulas to R's [plotmath expressions](http://stat.ethz.ch/R-manual/R-patched/library/grDevices/html/plotmath.html). Plotmath expressions are used to enter mathematical formulas and symbols to be rendered as text, axis labels, etc. throughout R's plotting system. I find plotmath expressions to be quite opaque and fiddly; LaTeX is a de-facto standard for mathematical expressions, so this script might be useful to others as well.

*Note that at the moment, this script is at very early stages. It /will/ fail for even very straightforward LaTeX formulas. It may improve in the future.*

## Usage
Clone this repository and source the script. The script's only dependence is  the stringr package.
```{r}
source("./latex2exp.r")
```

The `latex2exp` function takes a LaTeX string and returns a plotmath expression suitable for use in plotting, e.g.,

```{r, eval=FALSE}
latex2exp('\\alpha^\\beta')
```
(note it is *always* necessary to escape the backslash, hence the double backslash).

The return value of latex2exp can be used anywhere a plotmath expression is accepted, including plot labels, legends, and text. 

The following example shows plotting in base graphics:
```{r, warning=FALSE}
x <- seq(0, 4, length.out=100)
alpha <- 1:5

plot(x, xlim=c(0, 4), ylim=c(0, 10), xlab='x', ylab=latex2exp('\\alpha  x^\\alpha\\text{, where }\\alpha \\in \\text{1:5}'), type='n', main=latex2exp('\\text{Using }\\LaTeX 2Exp\\text{ for plotting in base graphics}'))

invisible(sapply(alpha, function(a) lines(x, a*x^a, col=a)))

legend('topleft', legend=latex2exp(sprintf("\\alpha = %d", alpha)), lwd=1, col=alpha)
```

This example shows plotting in [ggplot2](http://ggplot2.org):
```{r, warning=FALSE}
library(plyr)
x <- seq(0, 4, length.out=100)
alpha <- 1:5
data <- mdply(alpha, function(a, x) data.frame(v=a*x^a, x=x), x)

p <- ggplot(data, aes(x=x, y=v, color=X1)) +
    geom_line() + 
    ylab(latex2exp('\\alpha  x^\\alpha\\text{, where }\\alpha \\in 1\\ldots 5')) +
    ggtitle(latex2exp('\\text{Using }\\LaTeX 2Exp\\text{ for plotting in ggplot2}')) +
    coord_cartesian(ylim=c(-1, 10)) +
    guides(color=guide_legend(title=NULL)) +
    scale_color_discrete(labels=lapply(sprintf('\\alpha = %d', alpha), latex2exp)) # Note that ggplot2 legend labels must be lists of expressions, not vectors of expressions

print(p)
```

You can quickly test out what a translated LaTeX string would look like by using `plot`: 
```{r, fig.width=10, fig.height=2}
plot(latex2exp("\\text{A }\\LaTeX\\text{ formula: } \\frac{2hc^2}{\\lambda^5}  \\, \\frac{1}{e^{\\frac{hc}{\\lambda k_B T}} - 1}"), cex=2)
```

## "Supported" LaTeX
Only a subset of LaTeX is supported, and not 100% correctly. The following should be supported:
```{r, echo=TRUE}
print(noquote(latex2exp.supported()))
``` 
are supported. Their rendering depends on R's interpretation of the plotmath expression.

To render text with spaces or punctuation interspersed in the formula, embed it in `\\text{My text}`. 

A few examples:

```{r, fig.width=12, fig.height=7}
latex2exp_examples()
```

## FAQ
### It's not working even though my LaTeX string is correct
This script will get easily confused by even very simple LaTeX formulas (as I mentioned, it's a work in progress!). Please file a bug. 

### The translation is incorrect/Why is this not a proper R package?
The script, at the present stage, is just a quick hack to let me use LaTeX formulas in my scripts. 

At some point, I will rewrite this with more sound code and share it as an R package. Stay tuned, and feel free to file bugs.