---
title: "LaTeX2Exp"
author: "Stefano Meschiari"
output: 
  md_document:
    variant: markdown_github
---
```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=5)
```
# latex2exp
**latex2exp** is an R package that parses and converts LaTeX math formulas to R's [plotmath expressions](http://stat.ethz.ch/R-manual/R-patched/library/grDevices/html/plotmath.html). Plotmath expressions are used to enter mathematical formulas and symbols to be rendered as text, axis labels, etc. throughout R's plotting system. I find plotmath expressions to be quite opaque, while LaTeX is a de-facto standard for mathematical expressions, so this package might be useful to others as well.

## Installation 
Install this package from CRAN:
```{r, eval=FALSE}
install.packages('latex2exp')
```
You can also install from GitHub using [devtools](http://cran.r-project.org/web/packages/devtools/index.html):
```{r, eval=FALSE}
devtools::install_github('stefano-meschiari/latex2exp')
```

## Usage
```{r}
library(latex2exp)
```

The `TeX` function takes a LaTeX string and returns a plotmath expression suitable for use in plotting, e.g.,

```{r, eval=FALSE}
TeX('$\\gamma^\\alpha$')
```

Before R 4.0, it is necessary to escape the backslash within the string literal: therefore, one writes `'$\\gamma$'` rather than `'$\gamma$'` (the latter will cause an error).

After R 4.0, it is recommended to use the new raw string literal syntax (see `?Quotes`). The syntax looks like `r"(...)"`, where `...` can contain any character sequence, including `\`:

```{r, eval=FALSE}
TeX(r'($\gamma^\alpha$)')
```

The return value of `TeX()` can be used anywhere a plotmath expression is accepted, including plot labels, legends, and text. 

The following example shows plotting in base graphics:
```{r, warning=FALSE}
x <- seq(0, 4, length.out=100)
alpha <- 1:5

plot(x, xlim=c(0, 4), ylim=c(0, 10), 
     xlab='x', ylab=TeX(r'($\alpha  x^\alpha$, where $\alpha \in 1 \ldots 5$)'), 
     type='n', main=TeX(r'(Using $\LaTeX$ for plotting in base graphics!)', bold=TRUE, italic=TRUE))

invisible(sapply(alpha, function(a) lines(x, a*x^a, col=a)))

legend('topleft', legend=TeX(sprintf(r'($\alpha = %d$)', alpha)), 
       lwd=1, col=alpha)
```

This example shows plotting in [ggplot2](http://ggplot2.org):
```{r, warning=FALSE}
library(purrr)
library(ggplot2)
library(tibble)

x <- seq(0, 4, length.out=100)
alpha <- 1:5
data <- map_df(alpha, ~ tibble(v=.*x^., x=x, alpha=.))

p <- ggplot(data, aes(x=x, y=v, color=as.factor(alpha))) +
    geom_line() + 
    ylab(TeX(r'($\alpha  x^\alpha$, where $\alpha \in 1\ldots 5$)')) +
    ggtitle(TeX(r'(Using $\LaTeX$ for plotting in ggplot2. I $\heartsuit$ ggplot!)')) +
    coord_cartesian(ylim=c(-1, 10)) +
    guides(color=guide_legend(title=NULL)) +
    scale_color_discrete(labels=lapply(sprintf(r'($\alpha = %d$)', alpha), TeX)) 
    # Note that ggplot2 legend labels must be lists of expressions, not vectors of expressions

print(p)
```

You can quickly test out what a translated LaTeX string would look like by using `plot`: 
```{r, fig.height=2}
plot(TeX(r'(A $\LaTeX$ formula: $\frac{2hc^2}{\lambda^5} \, 
               \frac{1}{e^{\frac{hc}{\lambda k_B T}} - 1}$)'), cex=2)
```


## Syntax
Use
```{r, eval=FALSE}
TeX(r'(My string containing math $\sin(2\pi\lambda x)$)')
```
to build a plotmath expression, ready for use in plots. If the parser cannot build a correct plotmath expression, it will `stop()` and show the invalid plotmath expression built.

Add parameters `bold=TRUE` or `italic=TRUE` to make the entire label bold or italic:
```{r, eval=FALSE}
TeX(r'(My string containing math $\sin(2\pi\lambda x)$)', bold=TRUE, italic=TRUE)
```

```{r, eval=FALSE}
TeX('latexString', output=c('expression', 'character', 'ast'))
```
If the `output` option is equal to `character`, it will return the string representation of the expression (which could be converted into an expression using `parse(text=)`).

If the `output` option is equal to `ast`, it will return the tree built by the parser (this is only useful for debugging).

- - -

```{r, eval=FALSE}
latex2exp_examples()
```
will show a demo of the supported LaTeX syntax.

- - -
```{r, eval=FALSE}
latex2exp_supported(plot=FALSE)
```
returns a list of supported LaTeX. If `plot=TRUE`, a table of symbols will be plotted.


## Supported LaTeX
Formulas should go between dollar characters ($).

Only a subset of LaTeX is supported. Greek symbols (\\alpha, \\beta, etc.) and the usual operators (+, -, etc.) are supported. Their rendering depends on R's interpretation of the plotmath expression.

In addition, the following should be supported:
```{r, fig.width=8, fig.height=8}
latex2exp_supported(plot=TRUE)
``` 

A few examples:

```{r, fig.width=8, fig.height=7}
latex2exp_examples()
```

## Changes

### 0.5.0 [03/14/2021]
* Adds parameters `bold` and `italic` to `TeX()`. These can be used to make the
  entire expression bold or italic.
* Adds `\phantom{}` ([PR](https://github.com/stefano-meschiari/latex2exp/pull/22))

### 0.4.0 [08/29/2015]
* Deprecated the `latex2exp()` function; use `TeX()` instead.
* Added `\lbrack` and `\rbrack` to type left and right square brackets.

### 0.3.3 [08/11/2015]
Fixes bug #4 ("fix parsing of numbers"), where certain numbers inside formulas where not parsed correctly.

### 0.3.2 [07/28/2015]
Fixes bug #3 ("subscript and superscript style"). `latex2exp` now renders combined subscripts and superscripts correctly.

### 0.3.1 [07/02/2015]
Fixes bug #2 (white space causes unexpected behaviour). `latex2exp` should now be a bit more compliant with how LaTeX handles whitespace.

### 0.3.0 [06/30/2015]
`latex2exp` is now a proper package.

### 0.2.0 [06/29/2015]
Formulas must now be enclosed between dollar characters ($), as in LaTeX proper. Text does not need to be enclosed in \\text tags anymore.
