---
title: "LaTeX2Exp"
author: "Stefano Meschiari"
output: 
  md_document:
    variant: markdown_github
---
*latex2exp* is a super-hacky, quick and dirty, ugly R function to convert LaTeX math formulas to R's [plotmath expressions](http://stat.ethz.ch/R-manual/R-patched/library/grDevices/html/plotmath.html). Plotmath expressions are used to enter mathematical formulas and symbols to be rendered as text, axis labels, etc. throughout R's plotting system. I find plotmath expressions to be quite opaque and fiddly; LaTeX is a de-facto standard for mathematical expressions, so this script might be useful to others as well.

*Note that at the moment, this script is at very early stages; translation is done through a set of regexes. It /will/ fail for even very straightforward LaTeX formulas. It may improve in the future.*

## Usage
Clone this repository and source the script. The script's only dependence is  the stringr package.
```{r}
source("latex2exp.r")
```

The `latex2exp` function takes a LaTeX string and returns a plotmath expression suitable for use in plotting, e.g.,

```{r, eval=FALSE}
latex2exp('\\alpha^\\beta')
```
(note it is *always* necessary to escape the backslash, hence the double backslash).

The return value of latex2exp can be used anywhere a plotmath expression is accepted, including plot labels, legends, and text. For example:
```{r, warning=FALSE}
x <- seq(0, 4, length.out=100)
alpha <- 1:5
plot(x, xlim=c(0, 4), ylim=c(0, 10), xlab='x', ylab=latex2exp('\\alpha x^\\alpha,\\text{ where }\\alpha \\in \\text{1:5}'), type='n')
for (a in alpha)
  lines(x, a*x^a, col=a)
legend('topleft', legend=latex2exp(sprintf("\\alpha = %d", alpha)), lwd=1, col=alpha)
```

## "Supported" LaTeX
Only a subset of LaTeX is supported, and not 100% correctly. Greeks, exponentiation, subscript, and the LaTeX commands
```latex
\sum, \prod, \int, \cdot, \times, \pm, \eq, \neq, \geq, \leq, \partial, \approx, \sim, \propto, \equiv, \infty, \in, \notin, \sqrt, \left( and \right), \mathbf, \mathit, \mathrm, \textbf, \textit, \text, \frac, \dot, \widehat, \tilde, \hat, \underline
``` 
are supported. Their rendering depends on R's interpretation of the plotmath expression.

To render normal text interspersed in the text, use \\text{My text}.

A few examples:

```{r, fig.width=9, fig.height=7}
latex2exp.examples()
```

## FAQ
### It's not working even though my LaTeX string is correct
This script will get easily confused by even very simple LaTeX formulas (as I mentioned, it's a work in progress!). Please file a bug. 

### The translation is incorrect/Why is this not a proper R package?
The script, at the present stage, is just a quick hack to let me use LaTeX formulas in my scripts. Therefore, it currently does not handle lots and lots of correct LaTeX formulas. 

At some point, I will rewrite this with more sound code and share it as an R package. Stay tuned, and feel free to file bugs.