---
title: "LaTeX2Exp"
author: "Stefano Meschiari"
output: 
  md_document:
    variant: gfm
---
```{r, setup, include=FALSE}
knitr::opts_chunk$set(fig.width=7, fig.height=5)
```
<!-- badges: start -->
[![R-CMD-check](https://github.com/stefano-meschiari/latex2exp/workflows/R-CMD-check/badge.svg)](https://github.com/stefano-meschiari/latex2exp/actions)
<!-- badges: end -->

# latex2exp

**latex2exp** is an R package that parses and converts LaTeX to R's custom [plotmath expressions](http://stat.ethz.ch/R-manual/R-patched/library/grDevices/html/plotmath.html). 

Expressions returned by `latex2exp` can be used to create formatted text and mathematical formulas and symbols to be rendered as axis labels, annotations, legends, titles, etc. throughout R's plotting system. 

## Installation 
Install this package from CRAN:
```{r, eval=FALSE}
install.packages('latex2exp')
```

You can also install from GitHub using [devtools](http://cran.r-project.org/web/packages/devtools/index.html):
```{r, eval=FALSE}
devtools::install_github('stefano-meschiari/latex2exp')
```

```{r, child='vignettes/using-latex2exp.Rmd'}
```

```{r, child='markdown/supported-latex-child.Rmd'}
```

## Changes

```{r, child="NEWS.md"}
```